---
title: 'TLSTTS - 2019 Season Analysis'
author: Nick Criswell
date: "`r format(Sys.Date())`"
slug: tlst-ts-2019-season-analysis
output:
  blogdown::html_page:
    toc: true
categories:
  - fantasy
  - football
  - non-technical
  - R
tags: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.height = 5, 
                      fig.width = 7, 
                      fig.align= 'center')

library(jsonlite)       # Pulling down JSON data
library(tidyverse)      # Tools for data analysis
library(glue)           # Sticking strings together
library(RColorBrewer)   # Color Pals
library(plotly)         # Interactive graphics
library(DT)             # JS datatables package
library(ggthemes)       # For 538 theme
library(scales)         # Labels for ggplot scales
library(widgetframe)    # putting DTs in sites
```

```{r get_data, cache=TRUE}
source("C:/sleeper_fantasy/R/global.R")
 
# Get league info for 2019 season
s19_info <- F_get_lg_info(user_id = user_id, season = 2019)

# unload some information from this
lg_id <- s19_info$lg_id
lg_owners <- s19_info$user_vw0
lg_rules <- s19_info$lg_rules

#### Combining Data ####========================================================

# pull matchup and stat information
mu_out <- F_get_weekly_matchup(season = 2019, week = 1:4)
mu_view0 <- mu_out$mu_view0

stat_out <- F_get_weekly_player_stats(season = 2019, week = 1:4)
stat_sum0 <- stat_out$player_stats_sum0
stat_melt0 <- stat_out$stats_melt1

# combine the matchup and stat data
mu_view1 <- mu_view0 %>% 
  left_join(stat_sum0, 
            by = c("player_id" = "player_id", 
                   "week" = "week"))

# add in the owner name
mu_view2 <- mu_view1 %>% 
  left_join(lg_owners, 
            by = c("roster_id" = "roster_id"))

# add in player information
mu_view3 <- mu_view2 %>% 
  left_join(player_view0, 
            by = c("player_id" = "player_id"))

```


## Overview

This site will serve as a repository for analysis of the 2019 The League Season Thunderdome: The Sequel NFL Fantasy season. Our league is hosted on the fantastic [Sleeper](https://sleeper.app/) platform and data ~~are~~ ~~is~~ ~~are~~ (ok, fine) is pulled using the [Sleeper API](https://docs.sleeper.app/). The analysis is completed using the [`R`](https://cran.r-project.org/) programming environment and pushed out as a website via the [`blogdown` package](https://bookdown.org/yihui/blogdown/), utilizing the [Hugo](https://gohugo.io) static web framework. [Netlify](https://www.netlify.com/) is the hosting platform, and all code can be found on my [GitHub page](https://github.com/ncriswell). 

I'll try to make weekly updates to the site as the season goes along and will also be doing write-ups here and there that contain analyses that span multiple seasons, go back to a particular season, or are just something which might be interesting to look at. Go back to the [home page](https://agitated-jackson-c07a26.netlify.com/) to and look for "TLSTTS" in the post title for those ones. These posts will obviously focus on analyzing the league results through data; if you want analysis that focuses on analyzing the league through the lens of some kind of constantly shifting mythology involving a pantheon of several gods aped from various world traditions, see the weekly newsletter that Joe sends out. 

## Updates

I don't really want a ton of posts on this site just to accommodate an analysis that updates each week. So I am going to overwrite this post each week with new results. I will also be tweaking the commentary here and there to accommodate for the cumulative results. Below, I will detail when/if there are any new features. 

### Week 3 - September 28, 2019

* Initial release
* Data available through the third week of the season
* Analysis included for Winners and Losers, Round Robin, and Position Performance


### Week 4 - October 2, 2019

* Data available through the fourth week of the season
* Added transaction summary data and downloadable tables

## Winners and Losers

To orient ourselves, we'll start with a simple view of the win-loss record for all owners in the league. 

```{r, fig.align='center', fig.height=5, fig.width=7}
weekly_owner_scores0 <- mu_view3 %>% 
  group_by(week, display_name, starter, matchup_id) %>% 
  summarise(tot_points = sum(player_score, na.rm = TRUE)) %>% 
  ungroup()

# join this back onto 
weekly_owner_scores1 <- weekly_owner_scores0 %>% 
  left_join(weekly_owner_scores0, by = c("week", "matchup_id", "starter")) %>% 
  filter(display_name.x != display_name.y) %>% 
  mutate(Week = as.factor(week),
         display_name.x = as.factor(display_name.x)) %>% 
  select(Week, display_name.x, tot_points.x, tot_points.y, starter, matchup_id) %>% 
  rename(`Owner` = display_name.x, 
         `Points For` = tot_points.x, 
         `Points Against` = tot_points.y) %>% 
  arrange(Week) %>% 
  mutate(Outcome = case_when(`Points For` > `Points Against` ~ "Win", 
                             TRUE ~ "Loss")) %>% 
  group_by(Week, starter) %>% 
  mutate(`Week Median` = median(`Points For`), 
         `Week Mean` = mean(`Points For`)) %>% 
  ungroup()

# Get the number of wins and losses by owner
owner_wl0 <- weekly_owner_scores1 %>% 
  filter(starter == TRUE) %>% 
  group_by(Owner, Outcome) %>% 
  summarise(records = n()) %>% ungroup()

# Now flip this into a table that people can download
owner_wl_melt0 <- owner_wl0 %>% 
  spread("Outcome", "records", fill = 0) %>% 
  mutate(`Winning %` = round(Win / (Win + Loss), 4)) 

# Add to this the points for and against
owner_points_fa0 <- weekly_owner_scores1 %>% 
  filter(starter == TRUE) %>% 
  group_by(Owner) %>% 
  summarise(`Points For` = round(sum(`Points For`), 2), 
            `Points Against` = round(sum(`Points Against`), 2))

owner_wl_melt1 <- owner_wl_melt0 %>% 
  left_join(owner_points_fa0) %>% 
  arrange(desc(`Winning %`), desc(`Points For`))


```

```{r}
# make the datatable

datatable(owner_wl_melt1,
          class = 'cell-border stripe',
          rownames = FALSE, 
          caption = "TLST: TS 2019 NFL Season: Current Standings",
          filter = 'top',
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip', 
                         buttons = list('excel', "csv"),
                         pageLength = 14, autoWidth = TRUE)) %>% 
  formatPercentage("Winning %", 1)



```


```{r,  fig.align='center', fig.height=5, fig.width=7}

# graph it
ggplot(data = owner_wl0, aes(x = Owner, y = records, fill = Outcome)) + 
  geom_bar(stat="identity", position = "dodge", color = "black") + 
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() + 
  ggtitle("TLST: TS - 2019 Win-Loss Record") + 
  labs(caption = "Data pulled through Sleeper API https://docs.sleeper.app/") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text(),
        axis.text.x = element_text(size = 6),
        axis.title.x =element_blank(),
        legend.title = element_blank()) +
  guides(fill = guide_legend(nrow = 1, keyheight = 0.5))

```

Both Jon and I lost for the first time this week, erasing our perfect records. Max continues to maintain a perfect winning percentage despite being in the lower half of total scoring for the league. Well done, Max. Otto and Nate can't really catch a break here. 

If you are interested in how each player has performed league to date, check out the table below. 


```{r}
mu_dt <- mu_view3 %>% 
  mutate(Week = as.factor(as.numeric(week)),
         Score = round(player_score, 2)) %>% 
  select(Week, display_name, matchup_id, position, last_name, 
         first_name, Score, starter) %>% 
  mutate_at(c("display_name", "matchup_id", "position", "last_name", "first_name"), as.factor) %>% 
  rename(Owner = display_name, 
         Matchup = matchup_id,
         Last = last_name, 
         First = first_name, 
         Position = position,
         Starter = starter) 

datatable(mu_dt,
          class = 'cell-border stripe',
          rownames = FALSE, 
          caption = "TLST: TS 2019 NFL Season: Weekly Stats",
          filter = 'top',
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip', 
                         buttons = list('excel', "csv"),
                         pageLength = 10, autoWidth = TRUE))


```

```{r, fig.align='center', fig.height=5, fig.width=7}
ggplot(data = weekly_owner_scores1 %>% filter(starter == TRUE),
       aes(x = matchup_id, y = `Points For`, fill = Outcome, label = Owner)) + 
  geom_bar( position = "dodge", stat = "identity", colour = "black") + 
  geom_text(position = position_dodge(width = 1),
             angle = 0, size = 2) + 
  geom_hline(aes(yintercept = `Week Median`)) + 
  geom_hline(aes(yintercept = `Week Mean`), linetype = "dotted") +
  facet_wrap(~Week) +
  scale_fill_brewer(palette = "Set1") + 
  coord_flip() + 
  ggtitle("TLST: TS - 2019 Season Weekly Results",
          subtitle = "Teams are grouped by matchup. Solid line = week median; dotted line = week average.") + 
  labs(caption = "Data pulled through Sleeper API https://docs.sleeper.app/") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text(),
        axis.text.x = element_text(size = 6),
        axis.title.x =element_blank(),
        legend.title = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()) +
  guides(fill = guide_legend(nrow = 1, keyheight = 0.5))
  
```

That week four graph is something else, right? The Tony/Nate matchup and the Zach/Jeremy matchup represent the two lowest and two highest scorers of the week. But, only one of them is allowed to win. So, Zach is really getting screwed this week and Tony is getting away with one.

## Round Robin Analysis 

The prior graph shows weekly performance against the median and average. This brings up a question about how each owner performs against all other owners each week-not just how they do against the person they are playing. We can create a graph that will show for each week how many teams each owner *would have beat* if we had to play every other owner every week. This view is a handy way to find owners who are consistently benefiting/suffering from favorable/unfavorable matchups. Each "row" is a week. Colors are added to show the win/loss outcome, and the number of round robin wins is indicated with black text. 

```{r}
# get the times that each week a player would win against all other playes
rr_view0 <- weekly_owner_scores1 %>% 
  filter(starter == TRUE) %>% 
  mutate(Result = case_when(`Points For` > `Points Against` ~ "W", 
                            TRUE ~ "L")) %>% 
  group_by(Week) %>% 
  mutate(num_over = rank(`Points For`)) %>% 
  ungroup() %>% 
  mutate(`RR Wins` = num_over - 1) %>%
  group_by(Owner) %>% 
  mutate(`Total RR Wins` = sum(`RR Wins`)) %>% 
  ungroup()

ggplot(data = rr_view0, 
       aes(x = fct_reorder(Owner, `Total RR Wins`), 
           y = `Week`, 
           fill = `Result`, 
           label = `RR Wins`)) + 
  geom_tile(colour = "black") + 
  geom_text() +
  scale_fill_brewer(palette = "Set1") +
  ggtitle("TLST: TS - 2019 Round Robin Analysis",
          subtitle = "Graph is arranged so total round robin wins increases left to right") +
  labs(caption = "Data pulled through Sleeper API https://docs.sleeper.app/") +
  theme_fivethirtyeight() + 
  theme(axis.title.x =element_blank(),
        axis.text.x = element_text(angle = 90, size =8, hjust = 1), 
        legend.title = element_blank()) +
  guides(fill = guide_legend(nrow = 1, keyheight = 0.5))
  

```

This graph does a really good job of showing people who are benefiting from favorable matchupes and people who might have gotten robbed here and there. This graph now looks crazy as Tony's blue box in week four contains a 1 and Zach's red box contains a 12. Tony could have beat only one player in the league-the one he played. Zach would have beaten every player in the league except for one-the one he played. This contributes to Zach's being the overall points and round robin leader but only being at 0.500 winning percentage.   

## Position Performance

Let's see how each position contributes to points for each owner. A downloadable table for all players, for all weeks is below as are some summary graphs. 

```{r}
mu_dt <- mu_view3 %>% 
  mutate(Week = as.factor(as.numeric(week)),
         Score = round(player_score, 2)) %>% 
  select(Week, display_name, matchup_id, position, last_name, 
         first_name, Score, starter) %>% 
  mutate_at(c("display_name", "matchup_id", "position", "last_name", "first_name"), as.factor) %>% 
  rename(Owner = display_name, 
         Matchup = matchup_id,
         Last = last_name, 
         First = first_name, 
         Position = position,
         Starter = starter) 

datatable(mu_dt,
          class = 'cell-border stripe',
          rownames = FALSE, 
          caption = "TLST: TS 2019 NFL Season: Weekly Stats",
          filter = 'top',
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip', 
                         buttons = list('excel', "csv"),
                         pageLength = 10, autoWidth = TRUE))


```



```{r}

weekly_owner_scores_sum0 <- weekly_owner_scores1 %>% 
  filter(starter == TRUE) %>% 
  group_by(Week) %>% 
  mutate(Owner_Rank = rank(`Points For`)) %>% 
  select(Owner, Owner_Rank)

weekly_owner_position0 <- mu_view3 %>% 
  group_by(week, display_name, starter, matchup_id, position) %>% 
  summarise(position_points = sum(player_score, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(Week = as.factor(week), 
         display_name = as.factor(display_name)) %>% 
  rename(Owner = display_name,
         `Points For` = position_points) %>% 
  ungroup() %>% 
  left_join(weekly_owner_scores_sum0)
  
ggplot(data = weekly_owner_position0 %>% filter(starter == TRUE),
       aes(x = Owner,  y = `Points For`, 
           fill = position, label = round(`Points For`, 0))) + 
  geom_bar(stat = "identity", colour = "black") + 
  geom_text(position = position_stack(vjust = 0.5), size = 2) + 
  scale_fill_brewer(palette = "Set1", name = "Position") +
  facet_wrap(~Week) + 
  ggtitle("TLST: TS - 2019 Season Weekly Position Scores")  +
    labs(caption = "Data pulled through Sleeper API https://docs.sleeper.app/") +
  coord_flip() + 
  theme_fivethirtyeight() + 
  theme(legend.title = element_blank()) +
  guides(fill = guide_legend(nrow = 1, keyheight = 0.5))

```

If you like this view, I would say don't get used to it. This isn't going to be readable by the end of the season. The excess of purple and yellow in the plot above tells us pretty clearly that running back and wide receivers run the show in fantasy football. Unless you have the best QB in the league...

Another useful view will let us see the relative strength of our positions against the rest of the league in a simpler view. We can plot the average position score by owner for all positions and add indicators for league average. 

```{r}
owner_position_view0 <- mu_view3 %>% 
  filter(!is.na(position), starter == TRUE) %>% 
  group_by(display_name, position, week) %>% 
  summarise(tot_contribution = sum(player_score, na.rm = TRUE)) %>% 
  group_by(display_name, position) %>% 
  summarise(avg_contribution = mean(tot_contribution, na.rm = TRUE)) %>% 
  ungroup()

# overall averages
lg_position_view0 <- owner_position_view0 %>% 
  group_by(position) %>% 
  summarise(avg_contribution = mean(avg_contribution, na.rm = TRUE)) %>% 
  ungroup()

ggplot(data = owner_position_view0, 
       aes(x = display_name, y = avg_contribution)) +
  geom_hline(data = lg_position_view0, 
             aes(yintercept = avg_contribution, colour = position),
             show.legend = FALSE, size = 1.25) +
  geom_point(aes(colour = position), size = 5, alpha = 0.8) + 
  scale_colour_brewer(name = "Position", palette = "Set1") + 
  scale_y_continuous(name = "Average Weekly Position Contribution") + 
  scale_x_discrete(name = "") + 
  ggtitle("Average Weekly Contribution by Position", 
          subtitle = "Vertical lines represent league weekly average") + 
  labs(caption = "Data pulled through Sleeper API https://docs.sleeper.app/") +
  coord_flip() + 
  theme_fivethirtyeight() + 
  theme(legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 1, keyheight = 0.5))

```

The New England defense continues to outpace most of our quarterbacks and Dave's own RBs. And look out, ~~Jake~~ Pat Mahomes, Jackson is now the highest scoring QB in the league. Its too bad the rest of Otto's team isn't good at all. My RB and WR numbers are very, very average. I think that is going to be a problem if Pat has any more down weeks. Jon's RB core is beastly, which helps to shore up his less-beastly WRs. 

## Transaction Summary

The fantastic Sleeper API also had end points to track transactions. We can pull data on Adds, Drops, Waiver Claims, and Trades. Unfortunately the API doesn't have end points for proposed but rejected trades. Per the Sleeper API documentation, their servers couldn't handle the storage space required by Jeremy's trade proposals. 

Here is a summary of the transaction count by owner. 

```{r}

trans_res <- lapply(1:4, function(m) fromJSON(glue("https://api.sleeper.app/v1/league/", 
                           lg_id, 
                           "/transactions/", m))) 

# this object is too complicated to handle in normal ways. We need to do manipulation
#  on it by each element


trans_ref_view0 <- lapply(trans_res, 
                           function(m) select(m, transaction_id,  
                                              type, status, roster_ids)) %>% 
  bind_rows(.id = "week") %>% 
  as_tibble() %>% 
  unnest(roster_ids) %>% 
  distinct()

trans_adds_view <- lapply(trans_res, 
                          function(m) 
                            select(m, adds) %>% 
                            flatten() %>% 
                            bind_rows() %>% 
                            bind_cols(transaction_id = m$transaction_id) %>%  
                            gather(key, value, -transaction_id) %>% 
                            filter(!is.na(value)) %>% 
                            rename(add_player_id = key, 
                                   add_player_roster = value)) %>% 
  bind_rows(.id = "week") %>% 
  as_tibble() %>% 
  distinct()
  

trans_drops_view <- lapply(trans_res, 
                           function(m) 
                             select(m, drops) %>% 
                             flatten() %>% 
                             bind_rows() %>% 
                             bind_cols(transaction_id = m$transaction_id) %>%  
                             gather(key, value, -transaction_id) %>% 
                             filter(!is.na(value)) %>% 
                             rename(drop_player_id = key, 
                                    drop_player_roster = value)) %>% 
  bind_rows(.id = "week") %>% 
  as_tibble() %>% 
  distinct()

# combine all of this together
trans_view0 <- trans_ref_view0 %>% 
  left_join(trans_adds_view, by = c("week", "transaction_id", "roster_ids" = "add_player_roster")) %>% 
  left_join(trans_drops_view, by = c("week", "transaction_id", "roster_ids" = "drop_player_roster")) %>% 
  distinct() %>% 
  left_join(lg_owners, by = c("roster_ids" = "roster_id")) %>% 
  left_join(select(player_view0, player_name, position, player_id), 
            by = c("add_player_id" = "player_id")) %>% 
  rename(`Added Player` = player_name, 
         `Added Position` = position) %>% 
  left_join(select(player_view0, player_name, position, player_id), 
            by = c("drop_player_id" = "player_id")) %>% 
  rename(`Dropped Player` = player_name, 
         `Dropped Position` = position) %>% 
  select(-add_player_id, -drop_player_id, -user_id, -roster_ids) %>% 
  mutate_all(as.factor) %>% 
  mutate(type = str_to_title(type), 
         status = str_to_title(status)) %>% 
  rename(Owner = display_name, 
         Week = week, 
         Type = type, 
         Status = status, 
         Transaction_ID = transaction_id)
  
# get summary status on this
trans_sum <- trans_view0 %>% 
  filter(Status == "Complete") %>% 
  group_by(Type, Owner) %>% 
  summarise(`Number of Transactions`  = length(unique(Transaction_ID)))

ggplot(data = trans_sum, 
       aes(x = Owner, y  = `Number of Transactions`, fill = Type)) + 
  geom_bar(stat = "identity", position = "dodge", colour = "black") + 
  scale_fill_brewer(palette = "Set1") + 
  scale_y_continuous(name = "Number of Transactions") + 
  scale_x_discrete(name = "Owner") + 
  ggtitle("TLST: TS 2019 - Transaction Count by Owner") +
  labs(caption = "Data pulled through Sleeper API https://docs.sleeper.app/") +
  coord_flip() + 
  theme_fivethirtyeight() + 
  theme(legend.title = element_blank(),
        axis.title = element_text(),
        axis.title.y = element_blank()) +
  guides(color = guide_legend(nrow = 1, keyheight = 0.5))
  


```

We see that our league-leader isn't stressing too much about making moves. You can't really argue with his results. 

For those more inclined to see have a view of the transaction-level data, I have included two tables below. One is for non-trade transactions and the other one includes trades. I am not sure of the best way to show this. I kind of wanted to show transactions next to each other if they occurred at the same time (if you added a player and dropped another). But there are instances of adds without drops and drops without adds. Also, trades get really hard to show in this way since an uneven number of players can be involved from each team. I am open to suggestions on this one. 


```{r}

datatable(trans_view0 %>% filter(Type != "Trade"),
          class = 'cell-border stripe',
          rownames = FALSE, 
          caption = "TLST: TS 2019 NFL Season: Non-Trade Transactions",
          filter = 'top',
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip', 
                         buttons = list('excel', "csv"),
                         pageLength = 10, autoWidth = TRUE))



```

```{r}

# melt the data to show trade information
trade_view0 <- trans_view0 %>% 
  filter(Type == "Trade") %>% 
  distinct() %>% 
  select(-Type, -Status, -`Added Position`, -`Dropped Position`) %>% 
  gather("Add/Drop", "Player", -Week, -Transaction_ID, -Owner) %>% 
  distinct() 

datatable(trade_view0,
          class = 'cell-border stripe',
          rownames = FALSE, 
          caption = "TLST: TS 2019 NFL Season: Trades",
          filter = 'top',
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip', 
                         buttons = list('excel', "csv"),
                         pageLength = 10, autoWidth = TRUE))



```


In the future, I think it would be interesting to see how these transactions shake out over the course of a season. We'll be able to see who are the biggest winners and losers from trades and that kind of thing. I am already worried about my Sony Michele trade. He just had his best week of the season and Gore had a down week.  

